---
title: "677_Notes_Quarto"
author: "Joe"
format: html
editor: visual
---

---
title: "677 Final"
author: "Joe Griffiths"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following are notes for Chapter 5 of *Elementary Decision Making*, titled "Uncertainty due to Ignorance of the State of Nature".

# Libraries

```{r warning=FALSE}
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(tidyverse))
```

# Data

First, the data from the tables featured will be created.

Following page 120, the following variables/conditions exist:

States of nature: $\theta_1$ = "Today is a sunny day." $\theta_2$ = "Today is a rainy day."

Available actions: $\alpha_1$ = "Wear fair-weather outfit." $\alpha_2$ = "Wear outfit with raincoat." $\alpha_3$ = "Wear outfit with raincoat, boots, rain hat, and umbrella."

**Table 5.1**

|                      | $\alpha_1$ | $\alpha_2$ | $\alpha_3$ |
|----------------------|------------|------------|------------|
| $\theta_1$ (no rain) | 0          | 1          | 3          |
| $\theta_2$ (rain)    | 5          | 3          | 2          |

## Tables

```{r Original Tables}
## Mr. Nelson's Loss of Utility l(sigma,alpha) from (pg. 120,table 5.1)
sigma_one <- c(0,1,3)
sigma_two <- c(5,3,2)
Loss_table <- as.data.frame(t(data.frame(sigma_one,sigma_two))) |> 
  rename(a1 = V1, a2 = V2, a3 = V3)

## Weather indication probabilities Table (p121 table 5.2)
sigma_one <- c(0.6,0.25,0.15)
sigma_two <- c(0.2,0.3,0.5)
Weather_table <- as.data.frame(t(data.frame(sigma_one,sigma_two))) |> 
  rename(State_1 = V1, State_2 = V2, State_3 = V3)

## Strategy Table (p121 table 5.3)
Strategy <- as.character(1:27)
x_1 <- c(rep("a1",9),rep("a2",9),rep("a3",9))
x_2 <- rep(c(rep("a1",3), rep("a2",3),rep("a3",3)),3)
x_3 <- rep(c("a1","a2","a3"),9)
Strategy_Table <- data.frame(Strategy,x_1,x_2,x_3)

remove(sigma_one,sigma_two,x_1,x_2,x_3,Strategy)
```

```{r}

# dfs used to assemble action tables--------------------------------------
a1_count <- function(column){a <- str_count(column,"a1")}
a1_df <- as.data.frame(lapply(Strategy_Table[,2:4],a1_count))

a2_count <- function(column){b <- str_count(column,"a2")}
a2_df <- as.data.frame(lapply(Strategy_Table[,2:4],a2_count))

a3_count <- function(column){c <- str_count(column,"a3")}
a3_df <- as.data.frame(lapply(Strategy_Table[,2:4],a3_count))
# ------------------------------------------------------------------------

# setting up classes in order to do matrix multiplication
tw <- as.matrix(t(Weather_table))
a1_df <- as.matrix(a1_df)
a2_df <- as.matrix(a2_df)
a3_df <- as.matrix(a3_df)

# matrix multiplication
tA1 <- t(a1_df %*% tw)
tA2 <- t(a2_df %*% tw)
tA3 <- t(a3_df %*% tw)

EL_calc <- function(strategy){ ### for any strategy 1-27, calculates the EL for sit. 1 and 2
  Action_Probability_Table <- data.frame(p_action1 = tA1[,strategy],
                            p_action2 = tA2[,strategy],
                            p_action3 = tA3[,strategy])*Loss_table
  Action_Probability_Table$EL <- rowSums(Action_Probability_Table)
  EL <- Action_Probability_Table[,4]
  EL
}

Loss_of_Utility_Table <- as.data.frame(lapply(1:27, EL_calc))
Loss_of_Utility_Table_Trans <- as.data.frame(t(Loss_of_Utility_Table)) |> 
  rename(SigmaOne=V1,SigmaTwo=V2) |> 
  mutate(StrategyNumber = 1:27)
row.names(Loss_of_Utility_Table_Trans) <- 1:27

Plot_Loss <- ggplot(Loss_of_Utility_Table_Trans,aes(SigmaOne,SigmaTwo))+
  geom_point()+
  xlim(0,NA)+
  ylim(0,NA)+
  labs(title = "Expected loss of utility for strategy")+
  geom_text(label=rownames(Loss_of_Utility_Table_Trans),
            hjust = 0,
            vjust = 1.3,
            check_overlap = T)+
  theme_light()
Plot_Loss

```

# Book Exercises

These are exercises listed under Table 5.5 on pg 123.

## 5.1

Compute the action probabilities and losses for strategy $s_{10}$.

```{r 5.1}
#using function above
EL_calc(10)

strategy <- 10

#pasted from first part in function above
Action_Probability_Table <- data.frame(p_action1 = tA1[,strategy],
                            p_action2 = tA2[,strategy],
                            p_action3 = tA3[,strategy])*Loss_table
  
Action_Probability_Table
```

## 5.2

Use the table of random digits to select an observation according to the probabilities corresponding to $\theta_1$. Apply $s_{10}$ to this observation and list the corresponding action and loss. Repeat ten times and average the ten losses.

```{r 5.2}
#??
```

## 5.3

If Mr.Nelson replaced his weather meter by a crustal ball which told him the state of nature, $\theta$, reasonable use of this information would provide him with *ideal action probabilities*. Present this table of ideal action probabilities. These are called ideal because with ordinary information no strategy could do this well.

```{r 5.3}
#just rounded all the prob tables?
```
